---
title: "homework 5"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


```{r load_libraries}
library(tidyverse)
```

## problem 2

Importing the raw homicides dataset:
```{r, message=FALSE}
library (readr)

urlfile="https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"

homicide_raw <-read_csv(url(urlfile)) %>% 
  janitor::clean_names()
```

#### Raw dataset description:
The dataset contains data of criminal homicides over the past decade in 50 American cities. 

* The total number of observations  are **`r nrow(homicide_raw)`** and the total number of variables are **`r ncol(homicide_raw)`**
* It includes variables **`r colnames(homicide_raw)`**. 


```{r summarizing homicides}
homicide = homicide_raw %>%
unite('city_state',"city":"state", remove = FALSE)%>%
  select(-city,-state)%>%
   mutate (homi_type = case_when (disposition == "Closed without arrest" ~ "unsolved", 
                                  disposition == "Open/No arrest" ~ "unsolved", 
                                  disposition == "Closed by arrest" ~ "solved" )) %>%
  group_by(city_state) %>%
  filter (city_state != "Tulsa_AL") %>%
  count(homi_type) %>%
  spread(key = homi_type, value = n)%>%
  mutate (total = solved + unsolved) %>%
  select (-solved) 
  
```

#### Filtering and Estimating Homicide proportion for Baltimore:
```{r Baltimore homicides}
baltimore = homicide %>%
  filter(city_state == "Baltimore_MD")
```

```{r Baltimore proportion}
balt_prop = 
  prop.test(baltimore$unsolved, baltimore$total)%>%
  broom::tidy() %>%
  select (estimate, conf.low, conf.high)

  knitr::kable(balt_prop)
```

#### Estimating Homicide proportion for each city 
```{r proportions for each cities}
city_prop = homicide %>%
  mutate(city_prop = list(broom::tidy(prop.test(unsolved, total, conf.level=0.95)))) %>%
  unnest(city_prop)%>%
  select (city_state, estimate, conf.low, conf.high)%>%
  ungroup()

  knitr::kable(city_prop)
 
```
