---
title: "homework 5"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


```{r load_libraries}
library(tidyverse)
```



## Problem 1

The code chunk below imports the data in individual spreadsheets contained in `./data/zip_data/`. To do this, I create a dataframe that includes the list of all files in that directory and the complete path to each file. As a next step, I `map` over paths and import data using the `read_csv` function. Finally, I `unnest` the result of `map`.

```{r, message =FALSE}
full_df = 
  tibble(
    files = list.files("data/zip_data/"),
    path = str_c("data/zip_data/", files)
  ) %>% 
  mutate(data = map(path, read_csv)) %>% 
  unnest(cols = c(data))
```

The result of the previous code chunk isn't tidy -- data are wide rather than long, and some important variables are included as parts of others. The code chunk below tides the data using string manipulations on the file, converting from wide to long, and selecting relevant variables. 

```{r}
tidy_df = 
  full_df %>% 
  mutate(
    files = str_replace(files, ".csv", ""),
    group = str_sub(files, 1, 3)) %>% 
  pivot_longer(
    week_1:week_8,
    names_to = "week",
    values_to = "outcome",
    names_prefix = "week_") %>% 
  mutate(week = as.numeric(week)) %>% 
  select(group, subj = files, week, outcome)
```

Finally, the code chunk below creates a plot showing individual data, faceted by group. 

```{r}
tidy_df %>% 
  ggplot(aes(x = week, y = outcome, group = subj, color = group)) + 
  geom_point() + 
  geom_path() + 
  facet_grid(~group)
```

This plot suggests high within-subject correlation -- subjects who start above average end up above average, and those that start below average end up below average. Subjects in the control group generally don't change over time, but those in the experiment group increase their outcome in a roughly linear way.

## problem 2

Importing the raw homicides dataset:
```{r, message=FALSE}
library (readr)

urlfile="https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"

homicide_raw <-read_csv(url(urlfile)) %>% 
  janitor::clean_names()
```

#### Raw dataset description:
The dataset contains data of criminal homicides over the past decade in 50 American cities. 

* The total number of observations  are **`r nrow(homicide_raw)`** and the total number of variables are **`r ncol(homicide_raw)`**
* It includes variables **`r colnames(homicide_raw)`**. 


```{r summarizing homicides}
homicide = homicide_raw %>%
unite('city_state',"city":"state", remove = FALSE)%>%
  select(-city,-state)%>%
   mutate (homi_type = case_when (disposition == "Closed without arrest" ~ "unsolved", 
                                  disposition == "Open/No arrest" ~ "unsolved", 
                                  disposition == "Closed by arrest" ~ "solved" )) %>%
  group_by(city_state) %>%
  filter (city_state != "Tulsa_AL") %>%
  count(homi_type) %>%
  spread(key = homi_type, value = n)%>%
  mutate (total = solved + unsolved) %>%
  select (-solved) 
  
```

#### Filtering and Estimating Homicide proportion for Baltimore:
```{r Baltimore homicides}
baltimore = homicide %>%
  filter(city_state == "Baltimore_MD")
```

```{r Baltimore proportion}
balt_prop = 
  prop.test(baltimore$unsolved, baltimore$total)%>%
  broom::tidy() %>%
  select (estimate, conf.low, conf.high)

  knitr::kable(balt_prop)
```

#### Estimating Homicide proportion for each city 
```{r proportions for each cities}
city_prop = homicide %>%
  mutate(city_prop = list(broom::tidy(prop.test(unsolved, total, conf.level=0.95)))) %>%
  unnest(city_prop)%>%
  select (city_state, estimate, conf.low, conf.high)%>%
  ungroup()

  knitr::kable(city_prop)
 
```

#### Plotting for estimates and 95% CI for each city
```{r plotting proportions}
plot_city = city_prop %>%
  mutate (city_state = reorder (city_state, estimate)) %>%
  ggplot(aes (x = city_state, y = estimate)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) + 
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 60, hjust = 1)) + 
  labs (title = "Plot for proportion estimate and 95% CI for each city",
        x = "City and State",
        y = "Proportion estimate with 95% CI")
  
plot_city
```

### problem 3
Setting up a simulation

```{r}
sim_in = function(n = 30, mu = 0, sigma = 5) {
    x = rnorm(n, mean = mu, sd = sigma)
    test = t.test (x, mu = 0, conf.level = 0.95) %>%
      broom::tidy() %>%
    select(estimate, p.value)
    
    test
}


```
generating 5000 datasets and repeating simulations for all levels of mu

```{r}
sim_5000 = vector("list", 5000)
for(i in 1:5000) {
  sim_5000 [[i]]=sim_in()
}

sim_5000 = sim_5000 %>% 
  bind_rows()
  
```
```{r}
sim_all = function(set){
  mu_all = vector("list", 5000)
  for(i in 1:5000){
    mu_all[[i]]=sim_in(mu=set)
  }
  power = mu_all %>% 
    bind_rows()
  
  power
  }

```

```{r}
sim_any = 
  tibble(
    mu_val = c(0,1,2,3,4,5,6),
    estimates = map(mu_val, sim_all)
  ) %>% 
  unnest(estimates) %>% 
  mutate(
    reject_if = ifelse(p.value< 0.05, TRUE, FALSE)
  )
```

#### Creating plots 

```{r plot for power and mu}
power_plot = sim_any %>%
  group_by(mu_val) %>%
  summarise(power = sum(reject_if)/5000) %>%
  ggplot(aes(x = mu_val, y = power)) +
  geom_point (aes(color = mu_val)) +
  geom_line() + 
  labs (title = " Power vs True value of mu",
        x = "True value of Mu",
        y = "Power")
  power_plot
```

From the graph it can be observed that as the true value of μ increases, the power of the test also increases. We can conclude that the effect size increases, the power of the test to reject the null also increases.

```{r plots}
reject_avg = sim_any %>%
  group_by(mu_val) %>%
  filter (reject_if == TRUE) %>%
  summarise (mean_reject = mean(estimate))

all_avg = sim_any %>%
  group_by(mu_val) %>%
  summarise (mean_all = mean(estimate))

avg_combined = full_join(reject_avg, all_avg, by = "mu_val") %>%
  pivot_longer(
    mean_reject:mean_all,
    names_to = "samples",
    values_to = "avg_estimates"
  )

all_plot = all_avg %>%
  ggplot (aes(x= mu_val, y= mean_all)) +
  geom_point (alpha = 0.5) + 
  geom_line () +
labs (title = "True vs Average estimate of μ^ for all sample",
        x = "True value of μ",
        y = "Average estimate of μ^")
all_plot

reject_plot = reject_avg %>%
  ggplot (aes(x= mu_val, y= mean_reject)) +
  geom_point (alpha = 0.5) + 
  geom_line () +
labs (title = "True value vs Average estimate of μ^ when null was rejected",
        x = "True value of μ",
        y = "Average estimate of μ^ for Null-rejected")
reject_plot

overlay_plot = avg_combined %>%
  ggplot (aes(x= mu_val, y= avg_estimates, group = samples)) +
  geom_point (aes(color = samples), alpha = 0.5) + 
  geom_line (aes(color = samples)) +
labs (title = "Overlay Plot for true value of μ vs Average estimate of μ^ in all sample and rejected sample",
        x = "True value of μ",
        y = "Average estimate of μ^ for Null-rejected")
overlay_plot
```


The overlay plot shows that the true mean and average mean diverge for smaller values of mean and are similar when the mean value increases. This is because when the true mean is greater, there is higher proportion of rejection due to larger effect size. So, as avg estimate of μ ^ for null is rejected (Y axis) increases, it converges with the effect true value of μ (X axis) due to higher power to reject the null.

